% Copyright © 2013 Edward O'Callaghan. All Rights Reserved.

\section{Groups} % (fold)
\label{sec:groups}

\begin{defn}[Binary operation]
 A $\textbf{binary operation}$ on a set $\mathcal{X}$ is a map
 $\circ : \mathcal{X} \times \mathcal{X} \to \mathcal{X}'$.
 $\textbf{N.B.}$ that the binary operation need not be closed.
\end{defn}

\begin{defn}[Magma]
 A $\textbf{magma}$ is a set $\mathcal{M}$ equipped with a binary operation $\circ$ that is closed
 under the operation on $\mathcal{M}$. We denote the magma as the tuple pair $(\mathcal{M}, \circ)$.
\end{defn}


\begin{defn}[Semi-group]
 A $\textbf{semi-group}$ is a set $\mathcal{G}$ equipped with binary operation that is $\emph{associative}$.
 Hence, a semi-group is a magma where the operation is $\emph{associative}$;
 That is, given any $x,y,z \in \mathcal{G}$ then $x \circ (y \circ z) = (x \circ y) \circ z \in \mathcal{G}$.
 We denote the semi-group as the tuple pair $(\mathcal{G}, \circ)$, not to be confused with a magma from context.
\end{defn}

\begin{defn}[Monoid]
 A $\textbf{semi-group with idenitity}$ or, $\textbf{monoid}$ for short, is a semi-group $(\mathcal{G}, \circ)$
 with a unique identity element $e \in \mathcal{G}$ such that $x \circ e = x = e \circ x \, \forall x \in \mathcal{G}$
\end{defn}


\begin{proof}[Proof: unquieness of idenitity]
 Assume some other identity $e^{'}$ exists in $\mathcal{G}$ then, $e^{'} = e^{'} \circ e = e \circ e^{'} = e. \qedhere$
\end{proof}


\begin{exmp}
 Given $\mathcal{G} = \mathbb{N}$ with the binary law of composition $\circ$ to be defined as arithmetic addition $+$.
 Then, $(\mathbb{N}, +)$ forms a semi-group with identity $0$. Verify the axioms.
\end{exmp}


\begin{defn}[Group]
 A $\textbf{group}$ is a monoid where every element has an inverse. An abelian group is a group that is commutative.
\end{defn}

\begin{exmp}
 Given $\mathcal{G} = \mathbb{Z}$ with the binary law of composition $\circ$ to be defined as arithmetic addition $+$.
 Then, $(\mathbb{Z}, +)$ forms a semi-group with identity $0$. Verify the axioms.
\end{exmp}

\begin{question}
 Why does the set of naturals $\mathbb{N}$ not form a group under multiplication, however does form a monoid?
\end{question}

\begin{defn}[Group order]
	If a group $\mathcal{G}$ has $n$ finitely many elements the $\emph{order}$, denoted
	$\left| \mathcal{G} \right| = n$, is the number of elements of $\mathcal{G}$.
\end{defn}

\begin{defn}[Group element order]
	For a element $g$ in some group $\mathcal{G}$ the order of $g$ is defined to be
	the least positive integer $k$ such that $g^k = e$, where $e$ denotes the group identity,
	with respect to the groups law of composition. In symbols, $o(g)=k$. If no such $k$ exists
	then $g$ is said to have infinite order.
\end{defn}

\begin{rem}
	A non-trivial element, $g \neq e$, of finite order, $o(g)=k < \infty$, is called a
	\emph{torsion element} and for when $k=2$ it is called an \emph{involution}.
\end{rem}

\begin{thm}
	Every finite group of even order has a non-trivial involution. That is, for some group
	$\mathcal{G}$ where $|\mathcal{G}| = 2n < \infty$ we have that, there exists some
	non-trivial element $g \neq e$ in $\mathcal{G}$ such that $g^2 = e$.
\end{thm}

\begin{exmp}[Matrix Groups]
	Linear maps of vector spaces form groups that have characteristic properties. For some
	vector space $\mathcal{V}$ over some field $\F$ we may define the following groups taking
	matrix multiplication as the binary law of composition.
	\begin{enumerate}[i.)]
		\item The \emph{General Linear} group defined by,
			\[
				GL(\F) \doteq \{ M \in \mathcal{M} : det(M) \neq 0 \}.
			\]
		\item The \emph{Special Linear} group defined by,
			\[
				SL(\F) \doteq \{ M \in GL(\F) : \det(M) = 1 \}.
			\]
		\item The \emph{Orthogonal} group defined by,
			\[
				O(\R) \doteq \{ M \in GL(\R) : M^{T} M = I \}.
			\]
		\item The \emph{Special Orthogonal} group defined by,
			\[
				SO(\R) \doteq \{ M \in O(\R) : \det(M) = 1 \}.
			\]
		\item The \emph{Unitary} group defined by,
			\[
				U(\C) \doteq \{ M \in GL(\C) : M^{*} M = I \}.
			\]
		\item The \emph{Special Unitary} group defined by,
			\[
				SU(\C) \doteq \{ M \in U(\C) : \det(M) = 1 \}.
			\]
		\item The \emph{Symplectic} group is defined by,
			\begin{align*}
				Sp(2n,\F) & \doteq \{ M \in \mathcal{M}_{2n} : M^{T} \Omega M = \Omega \}
				\intertext{where $\Omega$ is skew-symmetric and defined as the block matrix,}
				\Omega & \doteq
				\begin{pmatrix}
					0 & I_n \\
					-I_n & 0
				\end{pmatrix}.
			\end{align*}
	\end{enumerate}
\end{exmp}

\begin{exmp}[Lorentz Group]
	The Lorentz group is defined as,
	\[
		\mathcal{L}(\R) \doteq
		\{ M \in \mathcal{M}_{2}(\R) : M^{T} C M = C \}
	\]
	where $C$ describes the Lorentz inner product with respect to the
	standard basis, i.e.
	\[
		C =
		\begin{pmatrix}
			1 & 0 \\
			0 & -1
		\end{pmatrix}.
	\]
\end{exmp}

\begin{defn}[Automorphism Group]
	Suppose $\mathcal{S}=(S,*)$ is some algebraic structure and $\mathbb{S}$ is the set
	of automorphism of $\mathcal{S}$. Then we may define the structure $(\mathbb{S}, \circ)$,
	where $\circ$ is defined as functional composition, as the \emph{group of automorphisms}
	of $\mathcal{S}$, denoted $Aut(\mathcal{S})$ or $\mathscr{A}(\mathcal{S})$. That is,
	\[
		\mathscr{A}(\mathcal{S}) \doteq (\mathbb{S}, \circ) \text{ where }
		\mathbb{S} \doteq
		\{ \phi : \mathcal{S} \to \mathcal{S}, \text{ where } \phi \text{ is a isomorphic}. \}
	\]
	For some algebraic structure $\mathcal{S}$ on set $S$.
\end{defn}


% Cyclic Group Section.
\input{cyclicgroups.tex}

\subsection{Permutations} % (fold)
\label{sec:permutations}
Take a finite set X with $|X|=n$, then the transformations of X are called $\textbf{permutations}$ of the
elements of X. In particular, the group of permutations of $X=\{ 1, 2, \cdots, n \}$ is a $\textbf{symmetric group}$,
denoted $S_n$, with $\textbf{order}$ $|S_n|=n!$. Thus, by taking any subgroup of $S_n$ we have a
$\textbf{permutation group}$. Also note that, for finite sets, $\emph{permutation}$ and $\emph{bijective maps}$
refer to the same operation, namely rearrangement of elements of X.
Another way is to consider, a group $\mathcal{G}$ and set X.
Then a group action is defined as a group homomorphism $\varphi$ from $\mathcal{G}$ to the symmetric group of X.
That is, the action $\varphi: \mathcal{G} \to S_n(X)$, assigns a permutation of X to each element of the group $\mathcal{G}$ in the following way:
\begin{itemize}
 \item From the identity element $e \in \mathcal{G}$ to the identity transformation $\Id{X}$ of X, that is, $\varphi : e \to \Id{X}$;
 \item A product of group homomorphisms $\varphi \circ \psi \in \mathcal{G}$ is then the composite of permutations given by $\varphi$ and $\psi$ in X.
\end{itemize}
Given that each element of $\mathcal{G}$ is represented as a permutation. Then a group action can also be consider as a permutation representation.

A permutation $\sigma \in S_n$ can be written,
\[
 \sigma =
 \begin{pmatrix}
  1 & 2 & \cdots & n \\
  a_1 & a_2 & \cdots & a_n
 \end{pmatrix}
 \, \, \,
 \text{where } a_1 = \sigma(1), a_2 = \sigma(2), \cdots .
\]

The identity permutation $\Id{n} \in S_n$ is simply,
\[
 \Id{n} =
 \begin{pmatrix}
  1 & 2 & \cdots & n \\
  1 & 2 & \cdots & n
 \end{pmatrix}
\].

Since $|S_n|=n!$ then the total number of ways n elements maybe permuted is $n!$.

Take any two permutations $\sigma,\pi \in S_n$ then composition is well defined as $\textbf{functional composition}$
as follows.

Given,
\[
 \sigma =
 \begin{pmatrix}
  1 & 2 & \cdots & n \\
  a_1 & a_2 & \cdots & a_n
 \end{pmatrix}
 \, \, and \, \,
 \pi =
 \begin{pmatrix}
  a_1 & a_2 & \cdots & a_n \\
  b_1 & b_2 & \cdots & b_n
 \end{pmatrix}
\]
then,

\begin{align*}
 \pi \circ \sigma &=
 \begin{pmatrix}
  1 & 2 & \cdots & n \\
  \pi(a_1) & \pi(a_2) & \cdots & \pi(a_n)
 \end{pmatrix}
 \\
 &=
 \begin{pmatrix}
  1 & 2 & \cdots & n \\
  b_1 & b_2 & \cdots & b_n
 \end{pmatrix}
\end{align*}
.

A inverse of any permutation $\sigma \in S_n$ is given by,
\[
 \sigma^{-1} =
 \begin{pmatrix}
  1 & 2 & \cdots & n \\
  a_1 & a_2 & \cdots & a_n
 \end{pmatrix}^{-1}
 =
 \begin{pmatrix}
  a_1 & a_2 & \cdots & a_n \\
  1 & 2 & \cdots & n
 \end{pmatrix}
\]

\subsection{Permutation parity}
Consider the algebraic structure:
\[
	\triangle_n (x_1, \dots , x_n) = \prod_{i<j} (x_i - x_j)
\]
TODO..


\subsection{Symmetric Group}
TODO FIX sections??

\begin{defn}[Dihedral group]
	The \emph{dihedral group} $\mathcal{D}_n$ is defined as the symmetries of a regular n-gon.
	The order $| \mathcal{D}_n | = 2n$ as there are $n$ rotations and $n$ reflections.
\end{defn}

\subsection{Group actions} % (fold)
\label{subsec:groupaction}

For any mathematical object (e.g. sets, groups, vector spaces) $X$ an isomorphism
of $X$ is a symmetry of $X$. The set of all isomorphisms of $X$, or symmetries
of $X$, form a group called the symmetry group of $X$, denoted $Sym(X)$.
More formally;

\begin{defn}[Group action]
	An $\emph{action}$ of a group $\mathcal{G}$ on a mathematical object $X$ is
	a mapping $\mathcal{G} \times X \to X$, defined by $(g,x) \mapsto g . x$
	satisfying:
	\begin{itemize}
			\item $e . x = x \, \forall x \in X$ and
			\item $(g h) . x = g . (h . x) \, \forall g,h \in \mathcal{G}, x \in X$.
	\end{itemize}
	That is, we have the ($\emph{left}$) $\mathcal{G}$-action on $X$ and denote this
	by $\mathcal{G} \acts X$.
\end{defn}

Notice that we may study properties of the symmetries of some mathematical object $X$
without reference to the structure of $X$ in particular.


% Subgroup Section.
\input{subgroups.tex}

% Group Homomorphism Section.
\input{grouphomomorphisms.tex}
\input{characters.tex}

% Cosets Section.
\input{cosets.tex}

\subsection{Factor (or Quotient) groups} % (fold)
\label{subsec:factorgroups}
Let $\mathcal{G}$ be a commutative group and consider a subgroup $\mathcal{H}$.
Then $\mathcal{H}$ determines an equivalence relation in $\mathcal{G}$ given by
\[
	x \sim x' \mbox{ iff } x - x' \in \mathcal{H}.
\]
..

\subsection{Non-commutative Groups} % (fold)
\label{sec:noncommutative-groups}
A common class of non-commutative groups are transformation groups.
Note:
\begin{defn}[Transformation]
 A bijective map $\varphi: X \to X$ is called a $\textbf{transformation}$ of X.
 \begin{note}
  The most trivial case is the $\emph{idenitity map} \, \Id{X}$ by $\Id{X}(x) = x, \, \forall x \in X$.
 \end{note}
\end{defn}
Hence, there exists a inverse $\varphi^{-1}$ of $\varphi$ such that $\varphi^{-1} \circ \varphi = \Id{X} = \varphi \circ \varphi^{-1}$.
Now, take two transformations of X, $\varphi$ and $\psi$, and let the product $\varphi \circ \psi$ be well defined.
Then the set of all transformations of X form the group $\textbf{Transf(X)}$.
Since, given $\varphi , \psi , \phi \in Transf(X)$ then we have associativity, $\varphi \circ (\psi \circ \phi) = (\varphi \circ \psi) \circ \phi$.
We have identity $e = \Id{X} \in Transf(X)$ and so, inverses $\forall \varphi \in Transf(X) \exists ! \varphi^{-1} : \varphi \circ \varphi^{-1} = e$.
Closure follows from the composition of two transformations $\varphi$ and $\psi$, since $(\varphi \circ \psi)^{-1} = \psi^{-1} \circ \varphi^{-1}$.

A transformation group is a type of group action which describes symmetries of objects. More abstractly,
since a group $\mathcal{G}$ is a category with a single object in which every morphism is bijective.
Then, a group action is a $\emph{forgetful functor}$ $\mathcal{F}$ from the group $\mathcal{G}$ in the category $\textbf{Grp}$
to the set category $\textbf{Set}$ that is, $\mathcal{F} : \mathcal{G} \to \textbf{Set}$.
